---
title: "Hospital Duration EDA"
author: "Hayoung Cheon"
date: "2025-02-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load the data
```{r}
setwd("C:/Users/oyunm/Desktop/SMU/DS-6372-Applied-Statistics/Project 1")
hospital <- read.csv("HospitalDurations.csv");

head(hospital)
summary(hospital)
str(hospital)

nrow(hospital)
```

## DATA Description
There are `113` observations in the data.
There are 12 variables in total. Except for hospital ID which doesn't have any meaning as a predictor, There are 9 numerical variables and 2 categorical variables.
The response variable is Lengh of stay, `Lgth.of.sty`.  
| Numerical Variable | Caegorical Variable |
|--------------------|---------------------|
| Length of Stay     | Medical School Affiliation(1=Yes, 2=No)  |
| Infecion risk      | Region(1=NE, 2=NC, 3=S, 4=W )  | 
| Routine culturing ratio|                  |
| Routine chest X-ray ratio|                  |
| Number of beds     |                  |
| Average Daily census|                  |
| Number of nurses   |                  |
| Available facilies |                  |



# Check for missing data
No missing DATA?
```{r}
library(naniar)
vis_miss(hospital)
```

# boxplots for each variables
```{r}
library(ggplot2)

# Create a loop that displays all variables of boxplots except for ID.
for (i in 2:12) {
  # Ensure the column index is valid
  if (i <= ncol(hospital)) {
    p <- ggplot(hospital, aes(x = "", y = hospital[[i]])) + 
      geom_boxplot() + 
      labs(y = colnames(hospital)[i], x = "") +
      ggtitle(paste("Boxplot of", colnames(hospital)[i])) +
      theme_minimal()
    
    print(p)
  }
}


```

# Scatter plot of Length of stay vs. all other numerical predictors.
```{r}
# Create a data set that dropped categorical variables and ID
hospital2 <- hospital[, -c(8, 9)]
summary(hospital2)

# Create a loop that displays all variables of boxplots except for ID.
for (i in 2:10) {
  # Ensure the column index is valid
  if (i <= ncol(hospital2)) {
    p <- ggplot(hospital2, aes(x = hospital2[[i]], y = hospital2$Lgth.of.Sty )) + 
      geom_point() + 
      labs(x = colnames(hospital2)[i], y = "Length of Stay") +
      ggtitle(paste("a scatterplot of", colnames(hospital2)[i])) +
      theme_minimal()
    
    print(p) 
  }
}

```

# Boxplot of Length of stay vs. categorical variables.
1) Med.Sc.Aff vs Lgth.of.sty: hospitals tha doesn't have medical school affiliation has a lower mean of length of stay. However, there are only 17 count for medical school affiliation and 96 that doesn't have medical school affiliation.
2) 
```{r}
#Med.Sc.Aff vs Lgth.of.Sty
ggplot(hospital, aes(x = as.factor(Med.Sc.Aff), y = Lgth.of.Sty, fill = as.factor(Med.Sc.Aff))) +
  geom_boxplot() +
  labs(x = "Medical School Affiliation", 
       y = "Length of Stay", 
       title = "A Boxplot of Medical School Affiliation vs Length of Stay") +
  theme_classic()


table(hospital$Med.Sc.Aff)

#Region vs Lgth.of.Sty
ggplot(hospital, aes(x = as.factor(Region), y = Lgth.of.Sty, fill = as.factor(Region))) +
  geom_boxplot() +
  labs(x = "Region", 
       y = "Length of Stay", 
       title = "A Boxplot of Region vs Length of Stay") +
  theme_classic()


table(hospital$Region)

```

# ggpair to check correlation and multicollinearity
```{r}
library(GGally)
ggpairs(hospital)
```

# Correlation matrix
The Infection Risk has the highest correlation with Length of stay, Region the next, Average Daily Census, then Rouine chest x-ray ratio.
Among variables, 
- `Avg.Pat`&`Avg.Nur`&`Pc.Ser.Fac` are highly correlated.
- `N.Beds` & `Avg.pat` & `Avg.Nur` are highly correlated.

```{r}
library(ggcorrplot)

# Compute correlation matrix
cor_matrix <- cor(hospital[, sapply(hospital, is.numeric)], use = "complete.obs")

# Plot using ggcorrplot
ggcorrplot(cor_matrix, 
           method = "square", 
           type = "lower",
           lab = TRUE, 
           lab_size = 3, 
           colors = c("blue", "white", "red"),
           title = "Correlation Matrix of Hospital Data",
           ggtheme = theme_minimal())

```

# Residual plots for each plot.
Inf.Risk has constant variance issue?
`Inf.Risk`
`R.CX.ray.Rat`
`Avg.Nur`

Influential point: 112 or 47

```{r}

for (i in 2:12) {
  # Ensure the column index is valid
  if (i <= ncol(hospital)) {
   model <- lm(Lgth.of.Sty ~ hospital[[i]], data=hospital)
   par(mfrow=c(2,2))
   p <- plot(model)
  
  }
}
```

# Find detail of Influential point
```{r}

```